---
title: "Playing with purrr!!"
author: "KH"
date: "May 22, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
In this mini session, we will be using `purrr` functions to automate some plots and run many models. 

#Install packages
```{r}
#install.packages("dslabs")
#install.packages("gapminder")
library(tidyverse)
library(dslabs)
library(gapminder)
library(tidyr)
library(magrittr)
library(broom)
library(gtools)
```

#Load dataset
```{r}
data("murders")

head(murders)
```

#add murder rate
```{r}
murders %<>%  mutate(rate = total / population * 100000)
```

#create plot function
```{r}
scatter_fun = function(x) {
  murders %>% filter(region == x) %>% 
  ggplot(aes(x = population/10^6, y = total, label = abb)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    xlab("Populations in millions (log scale)") +
    ylab("Total number of murders (log scale)") +
    ggtitle(paste0(x, " Region Gun Murders in 2010"))
}
```

#test
```{r}
scatter_fun("West")
```

#create plots
```{r}
regions <- unique(murders$region)

regions <- set_names(regions,regions)

region_plots <- map(regions, ~scatter_fun(.x) )
```

#example
```{r}
region_plots$South
```


#Purrr models 
```{r}
head(gapminder)

```

#make nested data
```{r}
nested_dat <- gapminder %>% 
  group_by(continent, country) %>% 
  nest()
```

#view nested data
```{r}
nested_dat$data[[1]]
```

```{r}
nested_dat %>% 
  arrange(continent, country)
```

#add models whoooo
```{r}
nested_dat %<>%
  mutate(model = map(data, ~lm(lifeExp ~ year, data = .)))
```

#another way
```{r}
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
```


```{r}
nested_dat %>%  mutate(model = map(data, country_model))
```

#create tidy function 
```{r}
tidy_glm <- function(x) {broom::tidy(x, conf.int = TRUE, effects = "fixed")}
```

#tidy it up
```{r}
nested_dat %<>%  mutate(tidy = map(model, tidy_glm)) 
```


#unnest results
```{r}
result <-  nested_dat %>%  unnest(tidy, .drop = T)
```
.drop = Should additional list columns be dropped? By default, unnest will drop them if unnesting the specified columns requires the rows to be duplicated.


#more tidying
```{r}
result %<>% mutate_at(vars(-continent, -country, -term), ~round(.x, 2))  %>%
  mutate(signif = stars.pval(p.value)) %>%  filter(term != "(Intercept)")

result
```
References:

+ Wickham, H., and Grolemund, G. (2017), R for Data Science, Sebastopol, CA: O'Reilly Media. 
+ https://emoriebeck.github.io/R-tutorials/purrr/
+ https://aosmith.rbind.io/2018/08/20/automating-exploratory-plots/
